---
title: "Cluster Analysis Script"
author: "Chris Cooper"
date: "2022-12-04"
output: html_document
---

The following page was used as a basis for how to add coloured labels to individual texts 

https://stackoverflow.com/questions/27485549/how-to-colour-the-labels-of-a-dendrogram-by-an-additional-factor-variable-in-r

Load the 'dendextend' library (install if not installed)
```{r}
library(dendextend)
```

Read the csv file with all the data
```{r}
yt <- read.csv('random_bnc200_dim.csv', row.names = 1)
head(yt)
tail(yt)
```

Specify the columns to be used for the cluster analysis 

- the dimension scores for Dimension 1 to Dimension 6
```{r}
data <- yt[,1:6]
head(data)
```

The clustering method used:

Agglomerative clustering, using the default method 'complete'
'The algorithm compares the farthest neighbours in all pairs 
of clusters and merges those clusters whose farthest neighbours
are the closest' (Levshina, 2015, p. 310). 

Ward.D2 might also be usable
```{r}
dend <- as.dendrogram(hclust(dist(data)))
```

Change the colors: YouTube '2'(red), BNC '4'(blue)

- these were numbers that I added to the dataframe before running this script
```{r}
labels_colors(dend) <- as.numeric(yt[,10][order.dendrogram(dend)])
```

Change the labels to 'Y' and 'B' (instead of filenames)

still can't read, but it's a bit cleaner
```{r}
labels(dend) <- paste(as.character(yt[,9])[order.dendrogram(dend)])
plot(dend)
```

load the 'cluster' library
```{r}
library(cluster)
```

use silhouette to test which number of clusters is optimal

I am checking from 2 to 15 clusters
```{r}
asw <- sapply(2:15, function(x) summary(silhouette(cutree(dend, 
              k = x), dist(data)))$avg.width)
asw
```
the highest asw is actually 3 (0.2929944), but it is very similar to 2 (0.2924318)

the 3rd cluster is very small and I'm only interested in the main BNC cluster
2 clusters would give me these texts


Next, give each text a label in df with cluster number (class-1 or class-2)

https://stackoverflow.com/questions/50856937/how-to-add-cluster-id-in-a-seperate-column-of-a-dataframe

- Show the labels
```{r}
solution <- cutree(dend, k = 2)
solution
```

Add the labels to the dataframe
```{r}
res <- cbind(yt, Class = factor(unname(solution), labels = c("class-1", "class-2")))
head(res)
```

Write the output to a csv file
```{r}
write.csv(res,'yt_bnc_clusters.csv')
```
from counting with pandas (in Python), I could see that 'Class-2' is the 
one I am looking for (with most BNC texts)

666 YouTube texts belong to the same cluster as BNC (1936 not is cluster)
